```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.align="center", fig.height=3, fig.width=4)
library(ggplot2)
theme_set(theme_bw(base_size=12))
library(dplyr)
```
## Homework 3
*Akshay Kumar Varanasi (av32826)*

**This homework is due on Feb. 12, 2019 at 4:00pm. Please submit as a PDF file on Canvas.**

In this homework, you are asked to evaluate two data sets and determine if they are tidy data sets. *We are referring to a very specific definition of "tidy", so if this term is unfamiliar to you, please review the lecture materials*.

**Problem 1: (2 pts)** The dataset `USAccDeaths` built into R contains accidental deaths in the US 1973-1978. You can run `?USAccDeaths` to learn more about this data set.
```{r}
USAccDeaths
```
Explain the variables present in this dataset. Using the variables in this dataset and the formal definition of tidy data that we learned in lecture, is this data set tidy? Explain why or why not.

Variables present in this dataset are Year, Month and Number. So instead of columns having name of the month and row names as years, there should be a column for months and a column for year.  Observations should be death of each person. Thus, it is not a tidy dataset.

The dataset `CO2` built into R contains data on the carbon dioxide uptake in grass plants. You can run `?CO2` to learn more about this data set.
```{r}
head(CO2)
```
Explain the variables present in this dataset. Using the variables in this dataset and the formal definition of tidy data that we learned in lecture, is this data set tidy? Explain why or why not.

Variables present in this are Plant, Type, Treatment, concentration and uptake. Each observation is of a plant. Therefore it satisfies the rules of tidy dataset.

**Problem 2: (5 pts)** Listed below are three examples of code that violate the rules in [section 2](https://style.tidyverse.org/syntax.html) of the tidyverse style guide. Which tidyverse style guidelines are violated in these example?
```{r, eval = FALSE}
iris %>% filter(Species=="versicolor") %>% head()
```

Violates 2.2.3 Infix operators. The operator == should be surrounded by spaces.

```{r, eval = FALSE}
iris[50,]
```

It violates 2.2.1 Commas guideline. It is always good to put a space after a comma, never before, just like in regular English.

```{r, eval = FALSE}
boxplot (len ~ dose, data = ToothGrowth, range = 1, width = c(2, 2, 2), varwidth = TRUE, notch = FALSE, outline = TRUE)
```

Here it violates 2.2.2 Parentheses. It is not good to put spaces inside or outside parentheses for regular function calls.

**Problem 3: (3 pts)** The `NCbirths` contains 1409 birth records from North Carolina in 2001. The column contents are as follows:

+ **Plural**: 1=single birth, 2=twins, 3=triplets.
+ **Sex**: Sex of the baby 1=male 2=female.
+ **MomAge**: Mother's age (in years).
+ **Weeks**: Completed weeks of gestation.
+ **Gained**: Weight gained during pregnancy (in pounds).
+ **BirthWeightGm**: Birth weight in grams.
+ **Low**: Indicator for low birth weight, 1=2500 grams or less, 0=otherwise.
+ **Premie**: Indicator for premature birth, 1=36 weeks or sooner, 0=otherwise.
+ **Marital**: Marital status: 0=married or 1=not married.

```{r}
NCbirths <- read.csv("http://wilkelab.org/classes/SDS348/data_sets/NCbirths.csv")
head(NCbirths)
```

For single births, what are the **max** completed weeks of gestation and the **mean** birth weight for babies that were born prematurely and for babies that were carried to term? State your answer in a sentence. **HINT:** Use the function `max()` to determine the maximum completed weeks of gestation.

```{r}

NCbirths %>% filter(Plural == "1") %>% group_by(Premie) %>% summarize(max = max(Weeks), mean = mean(BirthWeightGm))

```
For single birth babies, the max completed weeks of gestation and mean birth weight for babies is 45 weeks and 3430.724 gms in case of babies carried to term and 36 weeks and 2615.760 gms in case of premature birth. 
