```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.align = "center", fig.height = 5, fig.width = 6)
library(tidyverse)
theme_set(theme_bw(base_size = 12))
library(ggthemes)
```

## SDS 385 Report
*Akshay Kumar Varanasi (av32826)*

### Description of the data
Attribute Information:

1) ID number 2) Diagnosis (M = malignant, B = benign) 3-12)

Ten real-valued features are computed for each cell nucleus:

a) radius (mean of distances from center to points on the perimeter) b) texture (standard deviation of gray-scale values) c) perimeter d) area e) smoothness (local variation in radius lengths) f) compactness (perimeter^2 / area - 1.0) g) concavity (severity of concave portions of the contour) h) concave points (number of concave portions of the contour) i) symmetry j) fractal dimension ("coastline approximation" - 1)

Class distribution: 357 benign, 212 malignant

### Question
Find out which features are important for diagnosis of Breast Cancer?

### Answer
We find the important features for diagnosis by fitting a logistic regression model and based on the model we get, we decide which are important.

Call the necessary the libraries
```{r}
library(ggplot2)
library(plotROC)
```
Read the data from file
```{r}
cancer_data = read.csv("cancer_data.csv")
head(cancer_data)
```

Before we fit the model, we divide the data into training and testing data.
```{r}
train_fraction <- 0.5 # fraction of data for training purposes
set.seed(126) # set the seed to make the partition reproductible
train_size <- floor(train_fraction * nrow(cancer_data)) # number of observations in training set

train_indices <- sample(1:nrow(cancer_data), size = train_size)
train_data <- cancer_data[train_indices, ] # get training data
test_data <- cancer_data[-train_indices, ] # get test data
```

Now we fit the model with using all the features initially and then we remove one by one gradually till we get a reasonable model.
```{r}
glm_out <- glm(
diagnosis ~ radius_mean + texture_mean + perimeter_mean + area_mean + smoothness_mean + compactness_mean + concavity_mean + concave.points_mean + symmetry_mean + fractal_dimension_mean,
data = train_data,
family = binomial
) # family = binomial required for logistic regression
#summary(glm_out) # Not showing due to constraint on space.
```
 
After we successively remove predictors until only predictors with a p value less than 0.1 remain, we get the following model.
```{r}
# diagnosis ~ texture_mean + area_mean + concave.points_mean

glm_out <- glm(
diagnosis ~ texture_mean + area_mean + concave.points_mean,
data = train_data,
family = binomial
) # family = binomial required for logistic regression
summary(glm_out)
```

We create two different data frames one for training and one for test data. Now we see how good is this model on data by plotting the ROC curve.
```{r}
# results data frame for training data
df_train <- data.frame(
predictor = predict(glm_out, train_data),
known_truth = train_data$diagnosis,
data_name = "training"
)
# results data frame for test data
df_test <- data.frame(
predictor = predict(glm_out, test_data),
known_truth = test_data$diagnosis,
data_name = "test"
)
df_combined <- rbind(df_train, df_test)
ggplot(df_combined, aes(d = known_truth, m = predictor, color = data_name)) +
geom_roc(n.cuts = 0) +
scale_color_colorblind()
```

After looking at ROC curves we see that the model fits and predicts quite well on the data. Even the AUC values say the same thing that model is really good on both the training and test data. 
```{r}
p <- ggplot(df_combined, aes(d = known_truth, m = predictor, color = data_name)
) +
geom_roc(n.cuts = 0)
data_name <- unique(df_combined$data_name)
data_info <- data.frame(
data_name,
group = order(data_name)
)
left_join(data_info, calc_auc(p)) %>%
select(-group, -PANEL) %>%
arrange(desc(AUC))
```

So by fitting the model and testing it, we see that the model performs really good. This is because the diagnosis is dependent on the following three features: texture of the nuclei image , area of the nuclei , concave points. This is true because we identify cancer cells due to irregular nuclei size like large nuclei(area), different texture and the shape is not circular so cells have many concave points. Thus we found out the important features for diagnosis of breast cancer. 