```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.align = "center", fig.height = 5, fig.width = 6)
library(tidyverse)
theme_set(theme_bw(base_size = 12))
library(ggthemes)
```

## SDS 385 Report
*Akshay Kumar Varanasi (av32826)*

### Instructions

Write a brief (approximately 2--4 pages total, maximum is 5 pages) report on a data analysis topic of your choice. You should state a clear question, explain briefly the dataset you are using, and then provide an analysis to answer your question. All code required to perform the analysis must be included in the R Markdown document. Please do not include more than 2 figures in your analysis.

The knitted R Markdown document (as a PDF) *and* the raw R Markdown file (as .Rmd) as well as any required data files should be submitted to Canvas by 4:00pm on **April 16th, 2019**. These two documents will be graded jointly, so they must be consistent (as in, don't change the R Markdown file without also updating the knitted document!).

Grading comments:

1. This assignment will be graded pass/fail. See the course syllabus for details.

2. If you need more than 5 pages or more than 2 figures your project is too complex! You are not being graded on complexity. The goal is clarity and succinctness.

3. You are encouraged to work with data from your own research. If you would need extensive amounts of code to clean up the data and prepare for the analysis, then do this ahead of time, export the data to a clean csv file, and then work with that csv file here. Just explain in a few sentences what data the csv file contains and how it was generated.

4. Please don't include any extraneous code that does not contribute to answering the question.

5. The project instructions do not count towards the page limit. Feel free to delete them.

### Report

*Please add your report here.*


**Question**
Model to predict the sex of the baby

```{r}

library(tidyverse)
theme_set(theme_bw(base_size = 12)) # set default ggplot2 theme
library(ggthemes)
library(Stat2Data)
library(plotROC)
```


```{r}
data(NCbirths)
NCbirths <- na.omit(NCbirths)
#NCbirths <- read_csv("http://wilkelab.org/classes/SDS348/data_sets/NCbirths.csv")

# 
# NCbirths$Sex[NCbirths$Sex=='1']<-'0'
# NCbirths$Sex[NCbirths$Sex=='2']<-'1'
# NCbirths$Sex<-as.numeric(NCbirths$Sex)
head(NCbirths)
```

The column contents are as follows:

+ **Plural**: 1=single birth, 2=twins, 3=triplets.
+ **Sex**: sex of the baby 1=male 2=female.
+ **MomAge**: Mother's age (in years).
+ **Weeks**: Completed weeks of gestation.
+ **Gained**: Weight gained during pregnancy (in pounds).
+ **Smoke**: Mother is a smoker: 1=yes, 0=no.
+ **BirthWeightGm**: Birth weight in grams.
+ **Low**: Indicator for low birth weight, 1=2500 grams or less, 0=otherwise.
+ **Premie**: Indicator for premature birth, 1=36 weeks or sooner, 0=otherwise.
+ **Marital**: Marital status: 0=married or 1=not married.

```{r}

glm.out <- glm(Premie ~ Plural + BirthWeightGm + Gained, data = NCbirths, family = "binomial")

summary(glm.out)
```


```{r}

lr_data <- data.frame(predictor = glm.out$linear.predictors, prob = glm.out$fitted.values, NCbirths)

ggplot(lr_data, aes(x = predictor, y = prob, color = factor(Premie))) + geom_point() + scale_color_colorblind()


# make a data-frame with probabilities and premature births
pred_data <- data.frame(probability = glm.out$fitted.values, Premie = NCbirths$Premie)

# cutoff of 0.50
cutoff <- 0.50

# number of premature babies identified as premature (true positives)
pred_data %>%
  filter(probability > cutoff & Premie == 1) %>%
  tally() -> premie_true

# number of babies carried to term identified as babies carried to term (true negatives)
pred_data %>%
  filter(probability <= cutoff & Premie == 0) %>%
  tally() -> term_true

# total number of premature babies (known postives)
pred_data %>%
  filter(Premie == 1) %>%
  tally() -> premie_total
# total number of babies carried to term (known negatives)
pred_data %>%
  filter(Premie == 0) %>%
  tally() -> term_total

# true positive rate
tp <- premie_true$n / (premie_total$n)
# true negative rate
tn <- term_true$n / (term_total$n)
tp

# false positive rate (1 - true negative rate)
(1 - tn)



df1 <- data.frame(
  predictor = predict(glm.out, NCbirths),
  known_truth = NCbirths$Premie,
  model = "model1"
)
# plot

# the aesthetic names are not the most intuitive
# `d` (disease) holds the known truth
# `m` (marker) holds the predictor values
p<-ggplot(df1, aes(d = known_truth, m = predictor, color = model)) +
  geom_roc(n.cuts = 0) +
  coord_fixed() +
  scale_color_colorblind()
p
calc_auc(p)
```

```{r}
# K-means clustering
NCbirths %>% select(Plural,BirthWeightGm,Gained) %>% kmeans(centers = 2) -> km

df_kmeans <- data.frame(
  predictor = km$cluster,
  known_truth = NCbirths$Premie,
  model = "K-means"
)

p<-ggplot(df_kmeans, aes(d = known_truth, m = predictor, color = model)) +
  geom_roc(n.cuts = 0) +
  coord_fixed() +
  scale_color_colorblind()
p
calc_auc(p)
```

```{r}

```



```{r}
library(caret)

#train_fraction <- 0.4 # fraction of data for training purposes
set.seed(126) # set the seed to make the partition reproductible
#train_size <- floor(train_fraction * nrow(biopsy)) # number of observations in training set
#train_indices <- sample(1:nrow(biopsy), size = train_size)
#train_data <- biopsy[train_indices, ] # get training data
#test_data <- biopsy[-train_indices, ] # get test data

trainIndex <- createDataPartition(NCbirths$Sex, p = 0.5, 
                                  list = FALSE, 
                                  times = 1)

  
# Validation
# 10 fold repeated 3 times
train_control <- trainControl(method="repeatedcv", number=10, repeats=3)

# fix the parameters of the algorithm (k-fold)
#grid <- expand.grid(.fL=c(0), .usekernel=c(FALSE))

# bootstrap
# train_control <- trainControl(method="boot", number=100)
# define training control
#train_control <- trainControl(method="LOOCV")
# train the model
#model <- train(Species~., data=iris, trControl=train_control, method="nb")


head(NCbirths)

```


```{r}
NCbirths$Premie <- as.factor(NCbirths$Premie)
NCbirthsTrain <- NCbirths[ trainIndex,]
NCbirthsTest  <- NCbirths[-trainIndex,]


head(NCbirths)
glmFit1 <- train(Premie ~ Plural + BirthWeightGm + Gained , data = NCbirths, method="glm", family="binomial")
``` 

```{r}
#summary(glmFit1)
pred<-predict(glmFit1, newdata=NCbirths)
#dim(pred)
#accuracy <- table(pred, NCbirthsTest$Sex)
#sum(diag(accuracy))/sum(accuracy)

df2 <- data.frame(
  predictor = as.numeric(pred),
  known_truth = as.numeric(NCbirths$Premie),
  model = "model2"
)

q<-ggplot(df2, aes(d = known_truth, m = predictor, color = model)) +
  geom_roc(n.cuts = 0) +
  coord_fixed() +
  scale_color_colorblind()
q
calc_auc(q)
head(df2)
```


```{r}

## Generalized Boosted Regression Modeling (GBM)
gbmFit1 <- train(Premie ~ Plural + BirthWeightGm + Gained, data = NCbirths, method="gbm", verbose = FALSE)
pred2<-predict(gbmFit1, newdata=NCbirths)
#dim(pred)
#accuracy <- table(pred, NCbirthsTest$Sex)
#sum(diag(accuracy))/sum(accuracy)

df3 <- data.frame(
  predictor = as.numeric(pred2),
  known_truth = as.numeric(NCbirths$Premie),
  model = "model3"
)

r<-ggplot(df3, aes(d = known_truth, m = predictor, color = model)) +
  geom_roc(n.cuts = 0) +
  coord_fixed() +
  scale_color_colorblind()
r
calc_auc(r)

```

```{r}
## SVM kernel radial

svmfit <- train(Premie ~ Plural + BirthWeightGm + Gained, data = NCbirths, method="svmRadial", verbose = FALSE)
pred3<-predict(svmfit, newdata=NCbirths)
#dim(pred)
#accuracy <- table(pred, NCbirthsTest$Sex)
#sum(diag(accuracy))/sum(accuracy)

df4 <- data.frame(
  predictor = as.numeric(pred3),
  known_truth = as.numeric(NCbirths$Premie),
  model = "model4"
)

j<-ggplot(df4, aes(d = known_truth, m = predictor, color = model)) +
  geom_roc(n.cuts = 0) +
  coord_fixed() +
  scale_color_colorblind()
j
calc_auc(j)
```


```{r}
library(dplyr)
rffit<-randomForest(Premie ~ Plural + BirthWeightGm + Gained, data = NCbirths)
summary(rffit)
```

```{r}
library(MASS)
data(Aids2)
head(Aids2)
```


```{r}
# Don't take survival analysis
data("Melanoma")
head(Melanoma)
```

```{r}
library(ggplot2)
data(txhousing)
head(txhousing)
```

```{r}
data = read.csv("pd_speech_features.csv",header = TRUE)
head(data)
```

```{r}
#cancer_data = read.table("breast-cancer-wisconsin.data", fileEncoding="UTF-16", dec=",")
cancer_data = read.csv("breastcancer_data.csv")
head(cancer_data)

```

