```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.align="center", fig.height=3, fig.width=4)
library(ggplot2)
theme_set(theme_bw(base_size=12))
library(tidyr)
library(dplyr)
```
##Lab Worksheet 4

### Part 1: The dplyr pipe

The following questions in Part 1 are from Lab Worksheet 3. Answer these questions again, but this time use the dplyr pipe (`%>%`) in your answer.

**Problem 1:** *In an in-class exercise, we made the following plot of the Sitka dataset:*

```{r}
# download the sitka data set:
sitka <- read.csv("http://wilkelab.org/classes/SDS348/data_sets/sitka.csv")
head(sitka)
ggplot(sitka, aes(x=Time, y=size, group=tree)) + geom_line() + facet_wrap(~treat)
```

*Now modify the plot so that the line for each tree is colored according to the maximum size of the tree.* 
```{r}
sitka %>% 
  group_by(tree) %>%
  mutate(max_size = max(size)) %>%
  ggplot(aes(x=Time, y=size, group=tree)) + geom_line(aes(color = max_size)) + facet_wrap(~treat) 
    
```

**Problem 2:** *The package nycflights13 contains information about all flights departing from one of the NY City airports in 2013. In particular, the data table `flights` lists on-time departure and arrival information for 336,776 individual flights:*
```{r}
library(nycflights13)
flights
```

*We would like to collect some information about arrival delays of United Airlines (UA) flights. Do the following: pick all UA departures with non-zero arrival delay and calculate the mean arrival delay for each of the corresponding flight numbers. Which flight had the longest mean arrival delay and how long was that delay?*

```{r}
flights %>% 
  filter(carrier == "UA", arr_delay != 0) %>%
  group_by(flight) %>%
  summarize(mean_arr_delay = mean(arr_delay)) %>%
  filter(mean_arr_delay == max(mean_arr_delay))

```

Flight number 1510 has the longest mean arrival delay of 283 mins.

###Part 2: Combining data-frames with dplyr

**Problem 1:** *Invent two simple data sets that allow you explain the difference between the dplyr functions `left_join()` and `inner_join()`. Explain which features of your data sets affect the behavior of these two functions.*
```{r}
daily_max_temp <- read.csv(
  text = "Week,Max_temp
  Tue,12
  Wed,15
  Fri,14
  Sat,6
  Sun,8", stringsAsFactors = FALSE
)

daily_min_temp <- read.csv(
  text = "Week,Min_temp
  Mon,4
  Tue,6
  Thu,3
  Sat,2
  Sun,1", stringsAsFactors = FALSE
)

daily_temp <- left_join(daily_max_temp, daily_min_temp)
daily_temp
daily_temp <- inner_join(daily_max_temp, daily_min_temp)
daily_temp
```

Datasets created are daily max and min temperature. To see the difference some days have been ommitted in each dataset. When I used left join, since the left dataset is max temp, it sees if rows in left table is also there in right. If there are such rows, then data is joined. If not, then NA is displayed.   

In case of inner join, only common rows are displayed unlike the left join where all the rows in left table are displayed.

**Problem 2:** *I have split the sitka data set into two data-frames. First, look up the documentation for the `bind_rows` function. What does `bind_rows` do? Next, use `bind_rows` to combine `sitka1` and `sitka2` back into a single data-frame.*

Combines/joins the dataset rowwise instead of columnwise like left_join.
```{r}
sitka1 <- sitka[1:100,]
sitka2 <- sitka[101:395,]
head(sitka1)
head(sitka2)

sitka_combined <- bind_rows(sitka1, sitka2)
head(sitka_combined)
tail(sitka_combined)


```