```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.align = "center", fig.height = 3, fig.width = 4)
library(tidyverse)
theme_set(theme_bw(base_size = 12))
library(grid) # for `arrow()`
```
## Lab Worksheet 5

In 1898, Hermon Bumpus, an American biologist working at Brown University, collected data on one of the first examples of natural selection directly observed in nature. Immediately following a bad winter storm, he collected 136 English house sparrows, *Passer domesticus*, and brought them indoors. Of these birds, 64 had died during the storm, but 72 recovered and survived. By comparing measurements of physical traits, Bumpus demonstrated physical differences between the dead and living birds. He interpreted this finding as evidence for natural selection as a result of this storm:

```{r}
bumpus <- read_csv("http://wilkelab.org/classes/SDS348/data_sets/bumpus_full.csv")
head(bumpus)
```

The data set has three categorical variables (`Sex`, with levels `Male` and `Female`, `Age`, with levels `Adult` and `Young`, and `Survival`, with levels `Alive` and `Dead`) and nine numerical variables that hold various aspects of the birds' anatomy, such as wingspread, weight, etc.

*Question 1: Perform a PCA on the numerical columns of this data set. Then make three plots potting the data as PC2 vs. PC1, colored by (i) sex, (ii) age, (iii) survival.*

```{r}
# R code goes here.
bumpus %>% 
  select(-Sex, -Age, -Survival) %>%   # remove non numerical column
  scale() %>%            # scale to 0 mean and unit variance
  prcomp() ->            # do PCA
  bumpus_pca                    # store result as `pca`

# now display the results from the PCA analysis
bumpus_pca
```


```{r}
bumpus_pca$x
bumpus_pca_data <- data.frame(bumpus_pca$x, Sex = bumpus$Sex, Age = bumpus$Age, Survival = bumpus$Survival)
head(bumpus_pca_data)

library(ggthemes)
ggplot(bumpus_pca_data, aes(x = PC1, y = PC2, color = Sex)) + 
  geom_point() +
  scale_color_colorblind()


#
```

Higher PC2 values is generally for female while lower PC2 for male.

```{r}
ggplot(bumpus_pca_data, aes(x = PC1, y = PC2, color = Age)) + 
  geom_point() +
  scale_color_colorblind()
```

There is no dependance of PC on Age.

```{r}
ggplot(bumpus_pca_data, aes(x = PC1, y = PC2, color = Survival)) + 
  geom_point() +
  scale_color_colorblind()
```

There is no dependance of PC on Survival.

*Question 2: Now visualize the rotation matrix of the PCA obtained under Question 1.*

```{r}
# R code goes here.

bumpus_pca$rotation
# capture the rotation matrix in a data frame
bumpus_rotation_data <- data.frame(
  bumpus_pca$rotation, 
  variable = row.names(bumpus_pca$rotation)
)

# define a pleasing arrow style
arrow_style <- arrow(
  length = unit(0.05, "inches"),
  type = "closed"
)

# now plot, using geom_segment() for arrows and geom_text() for labels
ggplot(bumpus_rotation_data) + 
  geom_segment(aes(xend = PC1, yend = PC2), x = 0, y = 0, arrow = arrow_style) + 
  geom_text(aes(x = PC1, y = PC2, label = variable), hjust = 0, size = 3, color = "red") + 
  xlim(-1., 1.25) + 
  ylim(-1., 1.) +
  coord_fixed() # fix aspect ratio to 1:1

```

*Question 3: Given the four plots from Questions 1 and 2, how do you interpret PC1 and PC2? What does PC1 tell you about a data point? What does PC2 tell you about a data point?*

Closer the arrows more related they are like wingspread and weight are related i.e heavier birds have longer wingspread.

*Question 4: What percentage of the variation in the data does PC1 explain?*

```{r}
# R code goes here.

bumpus_percent <- 100*bumpus_pca$sdev^2 / sum(bumpus_pca$sdev^2)
bumpus_percent

perc_data <- data.frame(percent = bumpus_percent, PC = 1:length(bumpus_percent))
ggplot(perc_data, aes(x = PC, y = percent)) + 
  geom_col() + 
  geom_text(aes(label = round(percent, 2)), size = 4, vjust = -0.5) + 
  ylim(0, 80)

```

Almost 60 percent of the variation is explained by PC1.

*Question 5: Does the PCA suggest any specific physical characteristics for birds that survived? Consider only PC1 and PC2 for your answer.*

Answer goes here.

