```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.align="center", fig.height=4, fig.width=4)
```
## In-class worksheet 2
**Jan 24, 2019**

## 1. t test

We will try the t test on the built-in data set `PlantGrowth`. However, first we need to reformat the data set, which we do with the function `unstack()`. We store the reformatted data set in a variable `plants`:

```{r}
head(PlantGrowth)
plants <- unstack(PlantGrowth)
head(plants)
```

The data set contains plant growth yield (dry weight) under one control and two treatment conditions:
```{r}
boxplot(plants)
```

**Question:** Is the mean control weight significantly different from the mean weight under treatment 1? Is the mean weight under treatment 1 significantly different from the mean weight under treatment 2? Use the function `t.test()` to find out.

```{r}
# R code goes here.
t.test(plants$ctrl,plants$trt1)
?t.test
t.test(plants$trt1,plants$trt2)



```

## 2. Correlation

We will try the correlation test on the built-in data set `cars`. The data set contains the speed of cars and the distances taken to stop, measured in the 1920s:

```{r}
head(cars)
```

Is there a relationship between speed and stopping distance? Use the function `cor.test()` to find out. Then make a scatterplot of speed vs. stopping distance, using the function `plot()`.

```{r}
# R code goes here.
cor.test(cars$speed, cars$dist)
plot(cars$speed, cars$dist)

```
There is a significant positive relationship between a car’s speed and its stopping distance. The
correlation coefficient is 0.81, i.e., 66% of the variation in a car’s stopping distance is explained by the
car’s speed. (Remember, the square of the correlation coefficient, i.e. here 0.81^2=0.66, tells us the
amount of variation explained by the correlation.
## 3. Regression

We will do a regression analysis on the data set `cabbages` from the R package MASS. The data set contains the weight (`HeadWt`), vitamin C content (`VitC`), the cultivar (`Cult`), and the planting date (`Date`) for 60 cabbage heads:

```{r}
library(MASS) # load the MASS library to make the data set available
head(cabbages)
```

Use a multivariate regression to find out whether weight and cultivar have an effect on the vitamin C content. You will need to use the functions `lm()` and `summary()`.

```{r}
# R code goes here.
fit<-lm(VitC~Cult+HeadWt,data=cabbages)
summary(fit)

```
```{r}
anova(fit)
```
The conclusions remain unchanged. Both the cultivar and the weight have a significant effect on vitamin
C content
## 4. If this was easy

Look into the function `predict()`. Can you use it to estimate the vitamin C content of a c52 cultivar with a weight of 4? Can you use it to calculate the residuals of the regression model?

```{r}
d <- data.frame(Cult="c52", HeadWt=4) # make a data frame with 1 row
head(d)
predict(fit, d) # run predict on previously fitted model with new data frame
```
```{r}
residuals <- cabbages$VitC - predict(fit, cabbages)
plot(residuals, fit$residuals)
abline(0, 1)
```


