```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.align="center", fig.height=3, fig.width=4)
```
## In-class worksheet 5
**Feb 5, 2019**

## 1. Tidy data

Is the `iris` dataset tidy? Explain why or why not.
```{r}
head(iris)
```
This dataset is tidy because it satisfies all the three rules of tidy dataset.Measurement of different flowers is the observation and variables are the columns.

Is the `HairEyeColor` dataset tidy? Explain why or why not.
```{r}
HairEyeColor
```
This is not tidy as it is 2 dimensional table. This should have sex,hair color, eye color as columns.

## 2. Selecting rows and columns
All subsequent code will be based on the dplyr library, which is part of the tidyverse. So we first have to load this library:
```{r message=FALSE}
library(tidyverse)
```

Now, using the dplyr function `filter()`, pick all the rows in the `iris` dataset that pertain to species setosa, and store them in a new table called `iris_setosa`.
```{r }
iris_setosa<-filter(iris,Species=="setosa")
```

Pick all the rows in the `iris` dataset where species virginica has a sepal length > 7.
```{r }
iris_virginica<-filter(iris,Species=="virginica",Sepal.Length>7)
iris_virginica
```

Are there any cases in the `iris` dataset for which the ratio of sepal length to sepal width exceeds the ratio of petal length to petal width? Use `filter()` to find out.
```{r }
filter(iris,Sepal.Length/Sepal.Width>Petal.Length/Petal.Width)
```


Create a pared-down table which contains only data for species setosa and which only has the columns `Sepal.Length` and `Sepal.Width`. Store the result in a table called `iris_pared`.
```{r }
iris_pared<-select(iris_setosa,Sepal.Length,Sepal.Width)
head(iris_pared)
```

## 3. Creating new data, arranging

Using the function `mutate()`, create a new data column that holds the ratio of sepal length to sepal width. Store the resulting table in a variable called `iris_ratio`.
```{r }
iris_ratio<-mutate(iris,ratio=Sepal.Length/Sepal.Width)
head(iris_ratio)
```

Order the `iris_ratio` table by species name and by increasing values of sepal length-to-width ratio.
```{r}
arrange(iris_ratio,Species,ratio)
```

## 4. Grouping and summarizing

Calculate the mean and standard deviation of the sepal lengths for each species. Do this by first creating a table grouped by species, which you call `iris_grouped`. Then run `summarize()` on that table.
```{r }
iris_grouped<-group_by(iris,Species)
summarize(iris_grouped,mean=mean(Sepal.Length))

```

Use the function `n()` to count the number of observations for each species.
```{r }
iris_grouped<-group_by(iris,Species)
summarize(iris_grouped,num_obs=n())

```

For each species, calculate the percentage of cases with sepal length > 5.5.
```{r }
#iris_grouped<-group_by(iris,Species)
summarize(iris_grouped,percent=100*sum(Sepal.Length>5.5)/n())

```


## 5. If this was easy

Take the `iris_ratio` data set you have created and plot the distribution of sepal length-to-width ratios for the three species.

```{r }
library(ggplot2)
ggplot(iris_ratio,aes(x=ratio,fill=Species))+geom_density(alpha=0.5)

```

Now plot sepal length-to-width ratios vs. sepal lengths. Does it look like there is a relationship between the length-to-width ratios and the lengths? Does it matter whether you consider each species individually or all together? How could you find out?
```{r }
ggplot(iris_ratio, aes(y=ratio, x=Sepal.Length, color=Species)) + geom_point()

```
There seems to be an overall trend of increasing length-to-width ratio with increasing length, but it
seems that within each species there is little correlation between these values. We can check this by
running a correlation analysis for each species:

```{r}
# Setosa
setosa_ratio <- filter(iris_ratio, Species=='setosa')
cor.test(setosa_ratio$Sepal.Length, setosa_ratio$ratio)
```

```{r}
# Versicolor
versicolor_ratio <- filter(iris_ratio, Species=='versicolor')
cor.test(versicolor_ratio$Sepal.Length, versicolor_ratio$ratio)
```

```{r}
# Virginica
virginica_ratio <- filter(iris_ratio, Species=='virginica')
cor.test(virginica_ratio$Sepal.Length, virginica_ratio$ratio)
```

Virginica shows a significant correlation between sepal length and length-to-width ratio, the other two
species do not.